A decision support tool for the Home Health Care Routing and Scheduling Problem

### 2021-10-25 tex/Paper/main.tex
tidied up paper, made minor changes.

%----------------------------------------------------------------------------------%
\section{Literature review}
\label{sec:litreview}
\noindent The generalised growth in demand and availability of home health care services of recent years has resulted in a growth in academic interest in the related scheduling problem. While the problem might be addressed by drawing from the extensive vehicle routing problem (VRP) literature (see, for example, \cite{drexl2012} for a comprehensive review of VRPs and synchronisation constraints), HHCRSP has developed its own body of literature. In particular, there are two recent review articles, \cite{cisse2017} and \cite{fikar2017}. We refer the reader to them for a broader understanding of what is meant by HHCRSP and different solving approaches. Home Health Care and Home Care services across the world operate in different ways and, as such, the related scheduling problems have different objectives and constraints. For example, it might be possible to speed up the visits to some patients \citep{mosquera2019}, to allow nurses to share a minibus for part of their route, while walking the rest of it \citep{fikar2015}, for nurses to use public transport \citep{rest2016}, or to allow time preferences for patients \citep{decerle2018}. % This variety makes difficult comparing approaches and translating the findings from one application to another.

@@

% Exact methods
Several articles propose exact methods. These are usually based on a Mixed Integer Linear Program (MILP) formulation that is solved with the help of a commercial solver \citep{bredstrom2008,bachouch2011, mankowska2014, aithaddadene2016}. Typically, these formulations are three index formulations that involve two types of variables: $x_{ijk}$ binary variables that indicate if carer $k$ traverses the arc $i$ and $j$; and $t_{ik}$ real variables that determine at what time carer $k$ starts performing job $i$. These formulations are limited in practice and can only solve to optimality small problems (e.g. 45 patients in \cite{aithaddadene2016} or 10 in \cite{mankowska2014}). In order to address this, some researchers propose to decompose the problem in a column generation scheme. For example, \cite{rasmussen2012} tackles the HHCRSP problem in Denmark. They propose a branch-and-price scheme where generalised synchronisation constraints are relaxed and dealt with in the branching. % \correction{Only Rasmussen?}
 % \correction{Cite relevant papers here}. The ILP formulations work well for small problems, but are often not practical to solve for problems with a significant number of patients. %around 50

Often, in order to tackle larger instances, the MILPs are solved in a restricted manner and, while they do not guarantee optimality, they might provide feasible solutions suitable for day to day planning. This is often referred to as a matheuristic algorithm and can be done by reducing the size of the original problem by fixing some aspects and solving it again. One example of this technique is given in \cite{rasmussen2012} that also proposes clustering strategies to reduce the size of the problem, which allows them to solve problems with up to $150$ patients, although no longer proving optimality. Another is \cite{bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and carers in order to generate a series of restricted MILPs that can be solved in a short time.

Another approach, which is the one we use in this article, is to propose a metaheuristic algorithm independent of a mathematical solver. \cite{mankowska2014} and \cite{aithaddadene2016} propose an Adaptive Large Neighbourhood search and a hybrid GRASP combined with Iterated Local Search respectively. One advantage of the use of heuristic algorithms is that they do not require availability of a (often commercial) MILP solver, which is unlikely to be available to health care professionals.

@@

%----------------------------------------------------------------------------------%
\subsection{Skill mix}
\label{sub:skillmix}
\noindent Some jobs might require a specialised nurse to be performed, and nurses might only perform jobs that they are qualified to do. To model this, we define a function $S_1: N\times J \rightarrow \{0,1\}$, that takes value $1$ if nurse $i$ is skilled to perform job $j$, $S_1(i,j) = 1$ and $0$ otherwise. 
Naturally, for double service jobs, no single nurse is skilled to perform them $S_1(i,j) = 0, \forall i \in N, \forall j \in DS$. To account for this, we define a second function $S_2: N\times N\times J \rightarrow \{0,1\}$, that takes value $S_2(i_1,i_2,j) = 1$ if the combination of nurses $i_1$ and $i_2$ is sufficiently skilled to perform job $j$ and $0$ otherwise.

@@

%----------------------------------------------------------------------------------%
\subsection{Travel times}
\label{sub:traveltimes}
\noindent To describe travel times, for each nurse we define $N$ functions $T_i: J\cup\{k^i_0\} \times J\cup\{k^i_0\} \rightarrow \mathbb{R}$, $i \in N$, where $T_i(j_1, j_2)$ indicates the time it takes nurse $i$ to travel from job $j_1$ to job $j_2$, where $k^i_0$ refers to the start and end location of the nurse $i$. Each nurse might start from a different depot (typically, their own home), a fact that is accounted for in these functions. Note that with this definition we allow to have different travel times between jobs for different nurses, and this can be used to model, for example, different means of transport. In our observations, however, all nurses used their own vehicles, so we assume that $T_i$ returns driving times for all nurses and, in particular, that:
\begin{equation}
    T_{i_1}(j_1, j_2) = T_{i_2}(j_1, j_2),\ \forall i_1, i_2 \in N,\ \forall j_1, j_2 \in J. \label{eqn:traveltimequality}
\end{equation}

@@

%----------------------------------------------------------------------------------%
\subsection{Temporal constraints}
\label{sub:temporalconstraints}
\noindent Each job $j$ has a service time $s_j$ and a time window $t_j$ that is defined by its earliest start time, $e_j$ and the latest start time $l_j$. It is a hard constraint that jobs cannot start before their earliest start time:
\begin{align}
    t_j \geq e_j, & \forall j \in J, \label{eqn:jobstart}
\end{align}

@@

\paragraph{Total waiting time}\ \\
The total waiting time in a solution $\bar{W}$, calculated as:
\begin{equation}
     \bar{W} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{S}} w_{ij}.
\end{equation}

\paragraph{Total travel time}\ \\
If we denote by $TR(r_i)$ travel time associated with a route $r_i$, such that:
\begin{equation}
    TR(r_i) =  \sum_{h = 0}^{m_i - 1} T_i(k^i_h, k^i_{h + 1}),
\end{equation}
then the total travel time in $\mathcal{S}$, $\bar{T}$, is calculated as:
\begin{equation}
     \bar{T} = \sum_{i \in N}  TR(r_i).
\end{equation}

\paragraph{Total tardiness}\ \\
The total tardiness in a solution, $\bar{Z}$, is simply the sum of the tardiness across all jobs:
\begin{equation}
    \bar{Z} = \sum_{j \in J} z_j.
\end{equation}

\paragraph{Maximum tardiness}\ \\
The maximum tardiness in a solution, $Z_{max}$, is given by:
\begin{equation}
    Z_{max} = \max_{j \in J} z_j
\end{equation}

\paragraph{Total Overtime}\ \\
The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse represents the time by which their finish time exceeds their end of shift time, $en(i)$. If we denote the arrival of a nurse $i$ to its end point by $arr(i)$:
\begin{equation}
	arr(i) = st(i) + TR(r_i) + \sum_{k \in r_i} w_k + s_k,
\end{equation}
then the total overtime can be calculated as:
\begin{equation}
     \bar{O} = \sum_{i \in N} \max \{0, arr(i) - en(i)\}.
\end{equation}

\paragraph{Workload balance}\ \\
In order to produce practical solutions, maintaining a balance in the workload is fundamental. To achieve this, we quantify the minimum spare time on any nurse route in the schedule, $\bar{B}$. If $\bar{B}$ is minimised, we favour solutions in which the lowest spare time of any given nurse is as large as possible, hence removing imbalance when possible.
\begin{equation}
    \bar{B} = \min_{i \in N} \left\{\max \{0, en(i) - arr(i)\} \right\}.
\end{equation}

\paragraph{Total preference score}\ \\
The total preference score in the solution, $\bar{P}$, is computed as:
\begin{equation}
    \bar{P} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{R}} P(i,j).
\end{equation}
Finally, the full quality of the solution, $q(\mathcal{S})$, is then obtained as:
\begin{equation}
    q(\mathcal{S}) = \alpha_1 \bar{T} + % Travel time 
         \alpha_2 \bar{W} + % Waiting time 
         \alpha_3 \bar{Z} + % Tardiness 
         \alpha_4 \bar{O} + % Overtime 
         \alpha_5 \bar{B} + % Workload balance 
         \alpha_6 \bar{P} + % Preference score 
         \alpha_7 Z_{max}. \label{eqn:solutionquality}
\end{equation}

@@

\begin{table}[htbp!]
    \centering
    \caption{Weights proposed for the objective function of the DST}
    \begin{tabular}{ccc}
    \hline
    \textbf{Weight}  & \textbf{Value} & \textbf{Description} \\
    \hline
    $\alpha_1$      &   -1   &   Travel time \\
    $\alpha_2$      &   -1   &   Waiting time \\
    $\alpha_3$      &   -15   &   Tardiness \\
    $\alpha_4$      &   -15   &   Overtime \\   
    $\alpha_5$      &   $\frac{1}{2}$   &   Workload balance \\       
    $\alpha_6$      &   1   &   Preference score \\    
    $\alpha_7$      &   0   &   Maximum tardiness\\    
    \hline
    \end{tabular}
    \label{table:adequateweights}
\end{table}

@@

%----------------------------------------------------------------------------------%
\subsubsection{Type of Path Relinking}
\label{subsub:typepathrelinking}
\noindent Path relinking is performed at the end of the algorithm between all pairs of solutions in the pool. However, it is also possible to perform path relinking at intermediate steps of the algorithm too, each time a new solution is generated. Since this is potentially very time consuming, we have we have investigated the following three strategies for relinking intermediate solutions:
\begin{itemize}
    \item PR01 Perform path relinking between every generated solution and another solution selected randomly from the pool.
    \item PR02 Perform path relinking between every new solution and all solutions in the pool.
    \item PR03 Do not perform path relinking during the algorithm.
\end{itemize}

@@

%----------------------------------------------------------------------------------%
\subsection{Testing data}
\label{sub:testing}
\noindent For our experiments, we use three datasets. Two are from two similar problems in the literature, presented in \cite{mankowska2014} and \cite{aithaddadene2016}, and are used to evaluate the algorithm performance in Section~\ref{sub:algorithmperformance}. The third dataset is composed of randomly generated instances based on road distances in the city of Southampton. This dataset is used for tuning the algorithm parameters in Section~\ref{sub:parametertuning} and to explore different weights of the objective function in Section~\ref{sub:exploringobjfunction}. Details on the instances and the process for generating them are detailed in \ref{appendix:randomlygenerateddata}.

%----------------------------------------------------------------------------------%
\subsection{Parameter tuning}
\label{sub:parametertuning}
\noindent The GRASP algorithm is made up of several components, for each of which we have identified various potential configurations. %For example, we can choose whether to use or not some of the neighbourhoods described in Section~\ref{subseq:local_search}.
In order to select the best possible configuration automatically and in a statistically sound manner, we used the software package \texttt{irace} \citep{lopez2016} with a budget of 1000 runs. Table~\ref{table:parameters} shows a list of the hyper-parameters and solving strategies tested. In the first column we show the name of the hyper-parameter, in the second column a brief description, in the third column the type of variable we used for testing it, in the fourth column the values that were tested and in the fifth column the best value found by \texttt{irace}.

% "optparams=-do_twopt=0 -no_change_ls=1 -no_change_grasp=1 -pr_strategy=1 -pr_direction=1 -sols_in_pool=19 -grasp_dl=0.25 -grasp_dr=0.81 -rcl_strategy=2"
\begin{table}[h]
    \centering
    \caption{Parameters tested for configuring the GRASP algorithm, along with their type, values tested and the best configuration found by \texttt{irace}. Variable types used are binary (B), Categorical (C), Integer (I) and Real (R).}
    \begin{tabular}{cp{5.5cm}ccc}
    \hline
    \textbf{Parameter}& \textbf{Explanation}  & \textbf{Type} & \textbf{Tested} & \textbf{Best} \\
    \hline
    NE03                    & Use a local search neighbourhood to change nurse order & B & Yes, No               &  Yes  \\ % no_change_ls
    NE05                    &   Use the two opt neighbourhood & B & Yes, No                           & No   \\ % do_twopt
    NurseChange             &   Change nurse order between GRASP iterations & B & Yes, No                        &  Yes  \\ % no_change_grasp
    \texttt{pr\_strategy}   &  Path relinking strategy, as listed in Section~\ref{sub:pathrelinking}& C & 1, 2, 3   &   1 \\ %  
    \texttt{pr\_direction}  &   Path relinking direction: forward (1), backward (2), forward and backward (3), chosen randomly at each iteration (4).& C & 1, 2, 3, 4       &   1 \\ % performPathRelinking
    \texttt{pool\_size}     &  Number of solutions in pool & I & [1,100]                                            & 19    \\ % sols_in_pool
    \texttt{grasp\_dl}              & Minimum value of $\delta$ &   R & [0,0.5]                                     & 0.25  \\ % grasp_dl
    \texttt{grasp\_dr}              & Range of $\delta$ &   R & [0,1]                                               & 0.81\\ % grasp_dr
    \texttt{rcl\_strategy}          & Strategy for generating the reduced candidate list & C & 1,2          & 2 \\ % rcl_strategy    
    \hline
    \end{tabular}
    \label{table:parameters}
\end{table}

@@

%----------------------------------------------------------------------------------%
\subsection{Variations in the Literature}
\label{sub:literaturedata}
\noindent While the problems reported by \cite{mankowska2014} and \cite{aithaddadene2016} are very similar to the one we describe, there are slight variations in how the problem is modelled that need to considered. \cite{mankowska2014} describes the following objective function:
\begin{equation}
    \min \lambda_1 Dist + \lambda_2 Travel + \lambda_3 Travel^{max},
\end{equation}
where $Dist$ denotes the distance travelled by all nurses, $Travel$ denotes the total tardiness and $Travel^{max}$ denotes the maximum tardiness. For \cite{mankowska2014} tardiness refers to the time between the end of its time window and the actual start time of a job. Tardiness is not allowed for interdependent jobs, where the dependency gap is a hard constraint. Further, the synchronisation is simplified for jobs with two skills, where the skill with the lower index must always be performed earlier than the skill with the largest index.
Tardiness does not include overtime for the workers either, which is not mentioned in the model, so we have not set a limit for working hours of staff when solving these instances.

@@

\begin{table}[htbp!]
    \centering
    \caption{Weights used in the objective function to match those used in the methods from  \cite{aithaddadene2016} (AH) and \cite{mankowska2014} (MK). A weight of zero indicates that the measure is not considered, while a dash indicates that those measures that were considered infeasible, i.e. they contribute to $I(S)$.}
    \begin{tabular}{ccc}
    \hline
    \textbf{Weight}  & \textbf{Values for AH} & \textbf{Values for MK} \\
    \hline
    $\alpha_1$      &   $-0.3\times60$          &  $-\frac{1}{3}\times60$   \\
    $\alpha_2$      &   $0$                     &  $0$                      \\
    $\alpha_3$      &   -                       &  $-\frac{1}{3}\times60$   \\
    $\alpha_4$      &   -                       &  $0$   \\   
    $\alpha_5$      &   $0$                     &  $0$                      \\       
    $\alpha_6$      &   $-1$                    &  $0$                      \\    
    $\alpha_7$      &   -                       &  $-\frac{1}{3}\times60$             \\    
    \hline
    \end{tabular}
    \label{table:adaptedobjvalues}
\end{table}

@@

%----------------------------------------------------------------------------------%
\subsection{Algorithm Performance}
\label{sub:algorithmperformance}
\noindent While the problem description of our DST is slightly different from the examples seen in Section~\ref{sub:literaturedata}, we test the performance of GRASP algorithm, with the required modifications, in the instances provided by \cite{aithaddadene2016} and \cite{mankowska2014}.

For all the instances, we perform 5 runs and report the best and the average result. The full computational results and solving times are listed in the Appendix tables \ref{table:mkresults}, \ref{table:aithresults} and \ref{table:longerexperiments}.

In Table~\ref{table:grouptableaithcombined} we present a summary of the results found by our DST for the three instance groups presented in \cite{AitHaddadene2016}. The results are the average of the best and average results found for the instances of the three instance groups \emph{G1}, \emph{G2} and \emph{G3}. We have allowed a time of 5 seconds for instances in the group \emph{G1}, 5 minutes for those in \emph{G2} and three hours for the instances in \emph{G3}. See \ref{appendix:allowingextratime} for a note on how solution quality varies when increasing solving time.

@@

\noindent These results suggest that our DST can find better objective values within reasonable planning times. Further, we are able to replicate the 23 optimal solutions provided in the paper. The average results show that GRASP is also consistent across runs, finding the optimal solutions for all runs in group \emph{G1} and in most of the runs in group \emph{G2}.

Similarly, in Table~\ref{table:grouptablemkavgbest} we present the comparison of our DST against the performance of the AVNS algorithm presented in \cite{mankowska2014}.

>>>>

%----------------------------------------------------------------------------------%
\section{Literature review}
\label{sec:litreview}
\noindent The generalised growth in demand and availability of home health care services of recent years has resulted in a growth in academic interest in the related scheduling problem. While the problem might be addressed by drawing from the extensive vehicle routing problem (VRP) literature (see, for example, \cite{drexl2012} for a comprehensive review of VRPs and synchronisation constraints), HHCRSP has developed its own body of literature. In particular, there are recent review articles: \cite{cisse2017}, \cite{fikar2017}, \citet{dimascolo2021}, and \citet{grieco2020}. We refer the reader to them for a broader understanding of what is meant by HHCRSP and different solving approaches. Home Health Care and Home Care services across the world operate in different ways and, as such, the related scheduling problems have different objectives and constraints. For example, it might be possible to speed up the visits to some patients \citep{mosquera2019}, to allow nurses to share a minibus for part of their route, while walking the rest of it \citep{fikar2015}, for nurses to use public transport \citep{rest2016}, or to allow time preferences for patients \citep{decerle2018}. % This variety makes difficult comparing approaches and translating the findings from one application to another.

@@

% Exact methods
Several articles propose exact methods. These are usually based on a Mixed Integer Linear Program (MILP) formulation that is solved with the help of a commercial solver \citep{bredstrom2008,bachouch2011, mankowska2014, aithaddadene2016}. Typically, these formulations are three index formulations that involve two types of variables: $x_{ijk}$ binary variables that indicate if carer $k$ traverses the arc $i$ and $j$; and $t_{ik}$ real variables that determine at what time carer $k$ starts performing job $i$. These formulations are limited in practice and can only solve to optimality small problems (e.g. 45 patients in \cite{aithaddadene2016} or 10 in \cite{mankowska2014}). In order to address this, some researchers propose to decompose the problem in a column generation scheme. For example, \cite{rasmussen2012} tackles the HHCRSP problem in Denmark. They propose a branch-and-price scheme where generalised synchronisation constraints are relaxed and dealt with in the branching.% \correction{Only Rasmussen?}
 % \correction{Cite relevant papers here}. The ILP formulations work well for small problems, but are often not practical to solve for problems with a significant number of patients. %around 50

Often, in order to tackle larger instances, the MILPs are solved in a restricted manner and, while they do not guarantee optimality, they might provide feasible solutions suitable for day to day planning. This is often referred to as a matheuristic algorithm and can be done by reducing the size of the original problem by fixing some aspects and solving it again. One example of this technique is given in \cite{rasmussen2012} that also proposes clustering strategies to reduce the size of the problem, which allows them to solve problems with up to $150$ patients, although no longer proving optimality. Another is \cite{bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and carers in order to generate a series of restricted MILPs that can be solved in a short time. In addition, \citet{nikzad2021} considers a four-phase matheuristic that calls upon a variety of mixed integer models.

Another approach, which is the one we use in this article, is to propose a metaheuristic algorithm independent of a mathematical solver. Various metaheuristic approaches have been widely studied in the literature: \citet{manavizadeh2020} considers a simulated annealing (SA) algorithm, whilst \citet{nasir2020} investigates the effectiveness of a hybrid genetic algorithm (GA). Furthermore, a memetic algorithm is proposed by \citet{shahnejat2021}, and four hybrid metaheuristics including combinations of GA and variable neighbourhood search (VNS) and SA with local search (LS) are developed by \citet{liu2021}. \cite{mankowska2014} and \cite{aithaddadene2016} propose an Adaptive Large Neighbourhood search and a hybrid GRASP combined with Iterated Local Search respectively. One advantage of the use of heuristic algorithms is that they do not require availability of a (often commercial) MILP solver, which is unlikely to be available to health care professionals.

@@

%----------------------------------------------------------------------------------%
\subsection{Skill Mix}
\label{sub:skillmix}
\noindent Some jobs might require a specialised nurse to be performed, and nurses might only perform jobs that they are qualified to do. To model this, we define a function $S_1: N\times J \rightarrow \{0,1\}$, that takes value $1$ if nurse $i$ is skilled to perform job $j$, $S_1(i,j) = 1$ and $0$ otherwise. 
Naturally, for double service jobs, no single nurse is skilled to perform them $S_1(i,j) = 0, \forall i \in N, \forall j \in DS$. To account for this, we define a second function $S_2: N\times N\times J \rightarrow \{0,1\}$, that takes value $S_2(i_1,i_2,j) = 1$ if the combination of nurses $i_1$ and $i_2$ is sufficiently skilled to perform job $j$ and $0$ otherwise.

@@

%----------------------------------------------------------------------------------%
\subsection{Travel Times}
\label{sub:traveltimes}
\noindent To describe travel times, for each nurse we define $N$ functions $T_i: J\cup\{k^i_0\} \times J\cup\{k^i_0\} \rightarrow \mathbb{R}$, $i \in N$, where $T_i(j_1, j_2)$ indicates the time it takes nurse $i$ to travel from job $j_1$ to job $j_2$, where $k^i_0$ refers to the start and end location of the nurse $i$. Each nurse might start from a different depot (typically, their own home), a fact that is accounted for in these functions. Note that with this definition we allow to have different travel times between jobs for different nurses, and this can be used to model, for example, different means of transport. In our observations, however, all nurses used their own vehicles, so we assume that $T_i$ returns driving times for all nurses and, in particular, that:
\begin{equation}
    T_{i_1}(j_1, j_2) = T_{i_2}(j_1, j_2),\ \forall i_1, i_2 \in N,\ \forall j_1, j_2 \in J. \label{eqn:traveltimequality}
\end{equation}

@@

%----------------------------------------------------------------------------------%
\subsection{Temporal Constraints}
\label{sub:temporalconstraints}
\noindent Each job $j$ has a service time $s_j$ and a time window $t_j$ that is defined by its earliest start time, $e_j$ and the latest start time $l_j$. It is a hard constraint that jobs cannot start before their earliest start time:
\begin{align}
    t_j \geq e_j, & \forall j \in J, \label{eqn:jobstart}
\end{align}

@@

\subsubsection*{1. Total waiting time}
\noindent The total waiting time in a solution $\bar{W}$, calculated as:
\begin{equation}
     \bar{W} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{S}} w_{ij}.
\end{equation}

\subsubsection*{2. Total travel time}
\noindent If we denote by $TR(r_i)$ travel time associated with a route $r_i$, such that:
\begin{equation}
    TR(r_i) =  \sum_{h = 0}^{m_i - 1} T_i(k^i_h, k^i_{h + 1}),
\end{equation}
then the total travel time in $\mathcal{S}$, $\bar{T}$, is calculated as:
\begin{equation}
     \bar{T} = \sum_{i \in N}  TR(r_i).
\end{equation}

\subsubsection*{3. Total tardiness}
\noindent The total tardiness in a solution, $\bar{Z}$, is simply the sum of the tardiness across all jobs:
\begin{equation}
    \bar{Z} = \sum_{j \in J} z_j.
\end{equation}

\subsubsection*{4. Maximum tardiness}
\noindent The maximum tardiness in a solution, $Z_{max}$, is given by:
\begin{equation}
    Z_{max} = \max_{j \in J} z_j.
\end{equation}

\subsubsection*{5. Total overtime}
\noindent The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse represents the time by which their finish time exceeds their end of shift time, $en(i)$. If we denote the arrival of a nurse $i$ to its end point by $arr(i)$:
\begin{equation}
	arr(i) = st(i) + TR(r_i) + \sum_{k \in r_i} w_k + s_k,
\end{equation}
then the total overtime can be calculated as:
\begin{equation}
     \bar{O} = \sum_{i \in N} \max \{0, arr(i) - en(i)\}.
\end{equation}

\subsubsection*{6. Workload balance}
\noindent In order to produce practical solutions, maintaining a balance in the workload is fundamental. To achieve this, we quantify the minimum spare time on any nurse route in the schedule, $\bar{B}$. If $\bar{B}$ is minimised, we favour solutions in which the lowest spare time of any given nurse is as large as possible, hence removing imbalance when possible.
\begin{equation}
    \bar{B} = \min_{i \in N} \left\{\max \{0, en(i) - arr(i)\} \right\}.
\end{equation}

\subsubsection*{7. Total preference score}
\noindent The total preference score in the solution, $\bar{P}$, is computed as:
\begin{equation}
    \bar{P} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{R}} P(i,j).
\end{equation}
Finally, the full quality of the solution, $q(\mathcal{S})$, is then obtained as:
\begin{equation}
    q(\mathcal{S}) = \alpha_1 \bar{T} + % Travel time 
         \alpha_2 \bar{W} + % Waiting time 
         \alpha_3 \bar{Z} + % Tardiness 
         \alpha_4 \bar{O} + % Overtime 
         \alpha_5 \bar{B} + % Workload balance 
         \alpha_6 \bar{P} + % Preference score 
         \alpha_7 Z_{max}. \label{eqn:solutionquality}
\end{equation}

@@

\begin{table}[htbp!]
    \centering
    \caption{Weights proposed for the objective function of the DST}
    \begin{tabular}{ccc}
    \toprule
    \textbf{Weight}  & \textbf{Value} & \textbf{Description} \\
    \midrule
    $\alpha_1$      &   -1   &   Travel time \\
    $\alpha_2$      &   -1   &   Waiting time \\
    $\alpha_3$      &   -15   &   Tardiness \\
    $\alpha_4$      &   -15   &   Overtime \\   
    $\alpha_5$      &   $\frac{1}{2}$   &   Workload balance \\       
    $\alpha_6$      &   1   &   Preference score \\    
    $\alpha_7$      &   0   &   Maximum tardiness\\    
    \bottomrule
    \end{tabular}
    \label{table:adequateweights}
\end{table}

@@

%----------------------------------------------------------------------------------%
\subsubsection{Type of Path Relinking}
\label{subsub:typepathrelinking}
\noindent Path relinking is performed at the end of the algorithm between all pairs of solutions in the pool. However, it is also possible to perform path relinking at intermediate steps of the algorithm too, each time a new solution is generated. Since this is potentially very time consuming, we have we have investigated the following three strategies for relinking intermediate solutions:
\begin{itemize}
    \item PR01: Perform path relinking between every generated solution and another solution selected randomly from the pool.
    \item PR02: Perform path relinking between every new solution and all solutions in the pool.
    \item PR03: Do not perform path relinking during the algorithm.
\end{itemize}

@@

%----------------------------------------------------------------------------------%
\subsection{Testing Data}
\label{sub:testing}
\noindent For our experiments, we use three datasets. Two are from two similar problems in the literature, presented in \cite{mankowska2014} and \cite{aithaddadene2016}, and are used to evaluate the algorithm performance in Section~\ref{sub:algorithmperformance}. The third dataset is composed of randomly generated instances based on road distances in the city of Southampton. This dataset is used for tuning the algorithm parameters in Section~\ref{sub:parametertuning} and to explore different weights of the objective function in Section~\ref{sub:exploringobjfunction}. Details on the instances and the process for generating them are detailed in \ref{appendix:randomlygenerateddata}.

%----------------------------------------------------------------------------------%
\subsection{Parameter Tuning}
\label{sub:parametertuning}
\noindent The GRASP algorithm is made up of several components, for each of which we have identified various potential configurations. %For example, we can choose whether to use or not some of the neighbourhoods described in Section~\ref{subseq:local_search}.
In order to select the best possible configuration automatically and in a statistically sound manner, we used the software package \texttt{irace} \citep{lopez2016} with a budget of 1000 runs. Table~\ref{table:parameters} shows a list of the hyper-parameters and solving strategies tested. In the first column we show the name of the hyper-parameter, in the second column a brief description, in the third column the type of variable we used for testing it, in the fourth column the values that were tested and in the fifth column the best value found by \texttt{irace}.

% "optparams=-do_twopt=0 -no_change_ls=1 -no_change_grasp=1 -pr_strategy=1 -pr_direction=1 -sols_in_pool=19 -grasp_dl=0.25 -grasp_dr=0.81 -rcl_strategy=2"
\begin{table}[h]
    \centering
    \caption{Parameters tested for configuring the GRASP algorithm, along with their type, values tested and the best configuration found by \texttt{irace}. Variable types used are binary (B), Categorical (C), Integer (I) and Real (R).}
    \begin{tabular}{cp{5.5cm}ccc}
    \toprule
    \textbf{Parameter}& \textbf{Explanation}  & \textbf{Type} & \textbf{Tested} & \textbf{Best} \\
    \midrule
    NE03 & Use a local search neighbourhood to change nurse order & B & Yes, No & Yes \\ % no_change_ls
    NE05 & Use the two opt neighbourhood & B & Yes, No & No \\ % do_twopt
    NurseChange & Change nurse order between GRASP iterations & B & Yes, No & Yes \\ % no_change_grasp
    \texttt{pr\_strategy} & Path relinking strategy, as listed in Section~\ref{sub:pathrelinking} & C & 1, 2, 3 & 1 \\ %  
    \texttt{pr\_direction} & Path relinking direction: forward (1), backward (2), forward and backward (3), chosen randomly at each iteration (4) & C & 1, 2, 3, 4 & 1 \\ % performPathRelinking
    \texttt{pool\_size} & Number of solutions in pool & I & [1,100] & 19 \\ % sols_in_pool
    \texttt{grasp\_dl} & Minimum value of $\delta$ & R & [0,0.5] & 0.25 \\ % grasp_dl
    \texttt{grasp\_dr} & Range of $\delta$ & R & [0,1] & 0.81 \\ % grasp_dr
    \texttt{rcl\_strategy} & Strategy for generating the reduced candidate list & C & 1,2 & 2 \\ % rcl_strategy    
    \bottomrule
    \end{tabular}
    \label{table:parameters}
\end{table}

@@

%----------------------------------------------------------------------------------%
\subsection{Variations in the Literature}
\label{sub:literaturedata}
\noindent While the problems reported by \cite{mankowska2014} and \cite{aithaddadene2016} are very similar to the one we describe, there are slight variations in how the problem is modelled that need to be considered. \cite{mankowska2014} describes the following objective function:
\begin{equation}
    \min \lambda_1 Dist + \lambda_2 Travel + \lambda_3 Travel^{max},
\end{equation}
where $Dist$ denotes the distance travelled by all nurses, $Travel$ denotes the total tardiness and $Travel^{max}$ denotes the maximum tardiness. For \cite{mankowska2014} tardiness refers to the time between the end of its time window and the actual start time of a job. Tardiness is not allowed for interdependent jobs, where the dependency gap is a hard constraint. Furthermore, the synchronisation is simplified for jobs with two skills, where the skill with the lower index must always be performed earlier than the skill with the higher index.
Tardiness does not include overtime for the workers either, which is not mentioned in the model, so we have not set a limit for working hours of staff when solving these instances.

@@

\begin{table}[htbp!]
    \centering
    \caption{Weights used in the objective function to match those used in the methods from  \citet{aithaddadene2016} (AH) and \citet{mankowska2014} (MK). A weight of zero indicates that the measure is not considered, while a dash indicates that those measures that were considered infeasible, i.e. they contribute to $I(S)$.}
    \begin{tabular}{ccc}
    \toprule
    \textbf{Weight}  & \textbf{Values for AH} & \textbf{Values for MK} \\
    \midrule
    $\alpha_1$      &   $-0.3\times60$          &  $-\frac{1}{3}\times60$   \\
    $\alpha_2$      &   $0$                     &  $0$                      \\
    $\alpha_3$      &   -                       &  $-\frac{1}{3}\times60$   \\
    $\alpha_4$      &   -                       &  $0$   \\   
    $\alpha_5$      &   $0$                     &  $0$                      \\       
    $\alpha_6$      &   $-1$                    &  $0$                      \\    
    $\alpha_7$      &   -                       &  $-\frac{1}{3}\times60$             \\    
    \bottomrule
    \end{tabular}
    \label{table:adaptedobjvalues}
\end{table}

@@

%----------------------------------------------------------------------------------%
\subsection{Algorithm Performance}
\label{sub:algorithmperformance}
\noindent While the problem description of our DST is slightly different from the examples seen in Section~\ref{sub:literaturedata}, we test the performance of the GRASP algorithm, with the required modifications, in the instances provided by \cite{aithaddadene2016} and \cite{mankowska2014}. For all the instances, we perform 5 runs and report the best and the average result. The full computational results and solving times are listed in the Appendix tables \ref{table:mkresults}, \ref{table:aithresults}, and \ref{table:longerexperiments}.

In Table~\ref{table:grouptableaithcombined} we present a summary of the results found by our DST for the three instance groups presented in \cite{aithaddadene2016}. The results are the average of the best and average results found for the instances of the three instance groups \emph{G1}, \emph{G2} and \emph{G3}. We have allowed a time of 5 seconds for instances in the group \emph{G1}, 5 minutes for those in \emph{G2} and three hours for the instances in \emph{G3}. See \ref{appendix:allowingextratime} for a note on how solution quality varies when increasing solving time.

@@

\noindent These results suggest that our DST can find better objective values within reasonable planning times. In addition, we are able to replicate the 23 optimal solutions provided in the paper. The average results show that \comp{GRASP} is also consistent across runs, finding the optimal solutions for all runs in group \emph{G1} and in most of the runs in group \emph{G2}.

Similarly, in Table~\ref{table:grouptablemkavgbest} we present the comparison of our DST against the performance of the AVNS algorithm presented in \citet{mankowska2014}.

\begin{table}
	\centering
	\caption{Performance of the \comp{GRASP algorithm} (average of best and average results) compared to the AVNS algorithm presented in \citet{mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}. Figures in bold indicate the best results for each instance group. \comp{mk only 2dp, why?}}
	\begin{tabular}{c|c|ccc}
		\toprule
		        &  \emph{mk\_avns}    &   \multicolumn{3}{c}{GRASP} \\
		Group   &   Result    &   Best        &   Avg         &   Time limit\\
		\midrule
		\emph{A}    &   \textbf{225.6}  &   226.41  &   226.41      &   5 s\\
		\emph{B}    &   475.1   &   \textbf{408.25} &   408.31      &   30 s\\
		\emph{C}    &   713.6   &   \textbf{628.15} &   637.59      &   1 min\\
		\emph{D}    &   930.3   &   \textbf{798.80}  &   814.53      &   5 min\\
		\emph{E}    &   1064.7  &   \textbf{880.06} &   901.00       &   15 min\\
		\emph{F}    &   1588.0    &   \textbf{1405.68}&   1491.09     &   30 min\\
		\emph{G}    &   2161.2  &   \textbf{2042.10} &   2341.89     &   3 h\\
		\bottomrule
	\end{tabular}
	\label{table:grouptablemkavgbest}
\end{table}
